Automating G\"odel's Ontological Proof of God's Existence with Higher-order Automated Theorem Provers

### 2014-05-26 Papers/2014-ECAI/ECAI-51.tex
new final version

\begin{definition}\label{operators}
\HOML terms $s_\alpha$ are associated with type-raised \HOL terms
$\lift{s_\alpha}$ in the following way:
\begin{align*} {\lift{p_\alpha}} & = \hol{p}_{\lift{\alpha}}
  \\
  {\lift{X_\alpha}} & = \hol{X}_{\lift{\alpha}}
  \\
  {\lift{(s_{\alpha\ar\beta}\,t_{\alpha})}} & =
  {(\lift{s_{\alpha\ar\beta}}\,\lift{ t_{\alpha}})}
  \\
  {\lift{(\lam{X_\alpha} s_\beta)}} & = {(\hol{\lambda}
    \lift{X_\alpha} \lambdot \lift{s_\beta})}
  \\
  {\lift{(\neg_{o\ar o}\,s_{o})}} & =
  (\hol{\llift{\neg}_{\sigma\ar\sigma}}\,\lift{ s_{\alpha}})
  \\
  {\lift{(({\vee}_{o\ar o\ar o}\,s_{o})\,t_{o})}} & =
  ((\hol{\llift{\vee}_{\sigma\ar\sigma\ar\sigma}}\,\lift{
    s_{\alpha}})\,\lift{ t_{\alpha}})
  \\
  {\lift{((\forall_{(\alpha\ar o)\ar o}\,(\lam{X_\alpha} s_\beta)}} & =
  (\hol{\llift{\forall}_{(\alpha\ar\sigma)\ar\sigma}}\,(\hol{\lambda}
  \lift{X_\alpha} \lambdot \lift{s_\beta})
  \\
  {\lift{(\Box_{o\ar o}\,s_{o})}} & =
  (\hol{\llift{\Box}_{\sigma\ar\sigma}}\,\lift{ s_{\alpha}})
\end{align*}

@@

\bibliographystyle{ecai2014}
\bibliography{Bibliography}
\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\begin{definition}\label{operators}
\HOML terms $s_\alpha$ are associated with type-raised \HOL terms
$\lift{s_\alpha}$ in the following way:
\begin{align*} {\lift{p_\alpha}} & = \hol{p}_{\lift{\alpha}}
  \\
  {\lift{X_\alpha}} & = \hol{X}_{\lift{\alpha}}
  \\
  {\lift{(s_{\alpha\ar\beta}\,t_{\alpha})}} & =
  {(\lift{s_{\alpha\ar\beta}}\,\lift{ t_{\alpha}})}
  \\
  {\lift{(\lam{X_\alpha} s_\beta)}} & = {(\hol{\lambda}
    \lift{X_\alpha} \lambdot \lift{s_\beta})}
  \\
  {\lift{(\neg_{o\ar o}\,s_{o})}} & =
  (\hol{\llift{\neg}_{\sigma\ar\sigma}}\,\lift{ s_{\alpha}})
  \\
  {\lift{(({\vee}_{o\ar o\ar o}\,s_{o})\,t_{o})}} & =
  ((\hol{\llift{\vee}_{\sigma\ar\sigma\ar\sigma}}\,\lift{
    s_{\alpha}})\,\lift{ t_{\alpha}})
  \\
  {\lift{((\forall_{(\alpha\ar o)\ar o}\,(\lam{X_\alpha} s_\beta)}} & =
  (\hol{\llift{\forall}_{(\alpha\ar\sigma)\ar\sigma}}\,(\hol{\lambda}
  \lift{X_\alpha} \lambdot \lift{s_\beta})
  \\
  {\lift{(\Box_{o\ar o}\,s_{o})}} & =
  (\hol{\llift{\Box}_{\sigma\ar\sigma}}\,\lift{ s_{o}})
\end{align*}

@@

\bibliographystyle{ecai2014}
%\bibliography{Bibliography}
\begin{thebibliography}{10}

\bibitem{Adams}
R.M. Adams, `Introductory note to *1970', in {\em {Kurt G\"odel: Collected
  Works Vol. 3: Unpubl. Essays and Letters}}, Oxford Univ. Press, (1995).

\bibitem{AndersonGettings}
A.C. Anderson and M.~Gettings, `G\"odel ontological proof revisited', in {\em
  {G\"odel'96: Logical Foundations of Mathematics, Computer Science, and
  Physics: Lecture Notes in Logic 6}},  167--172, {Springer}, (1996).

\bibitem{anderson90:_some_emend_of_goedel_ontol_proof}
C.A. Anderson, `Some emendations of {G{\"o}del's} ontological proof', {\em
  Faith and Philosophy}, {\bf 7}(3), (1990).

\bibitem{Andrews:gmae72}
P.B. Andrews, `General models and extensionality', {\em Journal of Symbolic
  Logic}, {\bf 37}(2),  395--397, (1972).

\bibitem{andrewsSEP}
P.B. Andrews, `Church's type theory', in {\em The Stanford Encyclopedia of
  Philosophy}, ed., E.N. Zalta, spring 2014 edn., (2014).

\bibitem{C36}
C.~Benzm{\"u}ller, `{HOL} based universal reasoning', in {\em Handbook of the
  4th World Congress and School on Universal Logic}, ed., J.Y. Beziau~et al.,
  pp. 232--233, Rio de Janeiro, Brazil, (2013).

\bibitem{B5}
C.~Benzm{\"u}ller and D.~Miller, `Automation of higher-order logic', in {\em
  Handbook of the History of Logic, Volume 9 --- Logic and Computation},
  Elsevier, (2014).
\newblock Forthcoming; preliminary version available at
  {http://christoph-benzmueller.de/papers/B5.pdf}.

\bibitem{C34}
C.~Benzm{\"u}ller, J.~Otten, and Th. Raths, `Implementing and evaluating
  provers for first-order modal logics', in {\em Proc. of the 20th European
  Conference on Artificial Intelligence (ECAI)}, pp. 163--168, (2012).

\bibitem{B9}
C.~Benzm{\"u}ller and L.C. Paulson, `Exploring properties of normal multimodal
  logics in simple type theory with {LEO-II}', in {\em {Festschrift in Honor of
  {Peter B. Andrews} on His 70th Birthday}}, ed., C.~Benzm{\"u}ller~et al.,
  386--406, College Publications, (2008).

\bibitem{J23}
C.~Benzm{\"u}ller and L.C. Paulson, `Quantified multimodal logics in simple
  type theory', {\em Logica Universalis}, {\bf 7}(1),  7--20, (2013).

\bibitem{LEO-II}
C.~Benzm{\"u}ller, F.~Theiss, L.~Paulson, and A.~Fietzke, `{LEO-II} - a
  cooperative automatic theorem prover for higher-order logic', in {\em
  Proc.~of IJCAR 2008}, number 5195 in LNAI, pp. 162--170. Springer, (2008).

\bibitem{J30}
C.~Benzm{\"u}ller and B.~Woltzenlogel-Paleo, `Formalization, mechanization and
  automation of {G{\"o}del's} proof of {God's} existence', {\em
  arXiv:1308.4526}, (2013).

\bibitem{J28}
C.~Benzm\"uller and B.~Woltzenlogel-Paleo, `{G{\"o}del's God in Isabelle/HOL}',
  {\em Archive of Formal Proofs}, (2013).

\bibitem{W50}
C.~Benzm\"uller and B.~Woltzenlogel-Paleo, `G\"odel's {God} on the computer',
  in {\em Proceedings of the 10th International Workshop on the Implementation
  of Logics}, EPiC Series. EasyChair, (2013).
\newblock Invited abstract.

\bibitem{Coq}
Y.~Bertot and P.~Casteran, {\em {Interactive Theorem Proving and Program
  Development}}, Springer, 2004.

\bibitem{Nitpick}
J.C. Blanchette and T.~Nipkow, `Nitpick: A counterexample generator for
  higher-order logic based on a relational model finder', in {\em Proc. of ITP
  2010}, number 6172 in LNCS, pp. 131--146. Springer, (2010).

\bibitem{Satallax}
C.E. Brown, `Satallax: An automated higher-order prover', in {\em Proc. of
  IJCAR 2012}, number 7364 in LNAI, pp. 111 -- 117. Springer, (2012).

\bibitem{ContemporaryBibliography}
R.~Corazzon.
\newblock Contemporary~bibliography~on~ontological~arguments: {\scriptsize
  \url{http://www.ontology.co/biblio/ontological-proof-contemporary-biblio.htm}}.

\bibitem{Fitting}
M.~Fitting, {\em Types, Tableaux and G\"odel's God}, Kluwer, 2002.

\bibitem{fitting98}
M.~Fitting and R.L. Mendelsohn, {\em First-Order Modal Logic}, volume 277 of
  {\em Synthese Library}, Kluwer, 1998.

\bibitem{Gallin75}
D.~Gallin, {\em Intensional and Higher-Order Modal Logic}, North-Holland, 1975.

\bibitem{garbacz12:_prover_simpl_expal_away}
P.~Garbacz, `{PROVER9's} simplifications explained away', {\em Australasian
  Journal of Philosophy}, {\bf 90}(3),  585--592, (2012).

\bibitem{GoedelNotes}
K.~G\"odel, {\em Appx.A: Notes in Kurt G\"odel's Hand},  144--145.
\newblock In  \cite{sobel2004logic}, 2004.

\bibitem{Henkin50}
L.~Henkin, `Completeness in the theory of types', {\em Journal of Symbolic
  Logic}, {\bf 15}(2),  81--91, (1950).

\bibitem{homl}
R.~Muskens, `{Higher Order Modal Logic}', in {\em Handbook of Modal Logic},
  ed., P~Blackburn~et al.,  621--653, Elsevier, Dordrecht, (2006).

\bibitem{Isabelle}
T.~Nipkow, L.C. Paulson, and M.~Wenzel, {\em {Isabelle/HOL: A Proof Assistant
  for Higher-Order Logic}}, number 2283 in LNCS, Springer, 2002.

\bibitem{oppenheimera11}
P.E. Oppenheimera and E.N. Zalta, `A computationally-discovered simplification
  of the ontological argument', {\em Australasian Journal of Philosophy}, {\bf
  89}(2),  333--349, (2011).

\bibitem{rushby13}
J.~Rushby, `The ontological argument in {PVS}', in {\em Proc.~of CAV Workshop
  ``Fun With Formal Methods''}, St. Petersburg, Russia,, (2013).

\bibitem{Schulz:AICOM-2002}
S.~Schulz, `E -- a brainiac theorem prover', {\em {AI Communications}}, {\bf
  15}(2),  111--126, (2002).

\bibitem{ScottNotes}
D.~Scott, {\em Appx.B: Notes in Dana Scott's Hand},  145--146.
\newblock In  \cite{sobel2004logic}, 2004.

\bibitem{sobel2004logic}
J.H. Sobel, {\em Logic and Theism: Arguments for and Against Beliefs in God},
  Cambridge U. Press, 2004.

\bibitem{sutcliffe2009tptp}
G.~Sutcliffe, `The {TPTP} problem library and associated infrastructure', {\em
  Journal of Automated Reasoning}, {\bf 43}(4),  337--362, (2009).

\bibitem{J22}
G.~Sutcliffe and C.~Benzm{\"u}ller, `Automated reasoning in higher-order logic
  using the {TPTP THF} infrastructure.', {\em Journal of Formalized Reasoning},
  {\bf 3}(1),  1--27, (2010).

\bibitem{J29}
B.~Woltzenlogel-Paleo and C.~Benzm{\"u}ller, `Automated verification and
  reconstruction of {G\"odel's} proof of {God's} existence', {\em OCG J.},
  (2013).

\end{thebibliography}

\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
